---
title: "Write to Specified Location"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # Set WD to Root
here::i_am("dev/openxl-writeLoc.Rmd")
library(tidyverse)
library(here)
library(openxlsx)
source(here("my_R/mkdata.R"))
library(testthat)
```

### Script

```{r}
usethis::use_r("xl-loc")
usethis::use_r("xl-write")
```



### Write Excel Grid Vertical

write.xlsx_gV

args: 
`...` = list of DFs
file = file
asTable
overwrite
writeData_args

```{r}
lbdoc::write.xlsx_gV(tab1 = list(iris[1:5, ], mtcars[1:5, ]),
                     tab2 = list(Orange[1:3, ], iris[1:2, ]),
                     file = path_excel("Excel_VGridTest1.xlsx"), overwrite = T
                     #writeData_args = list(borderColour = "black"),
                     )
```



```{r write.xlsx_gV}
write.xlsx_gV <- function(..., 
                          file,
                          overwrite = FALSE,
                          startCol = 1,
                          startRow = 1, 
                          gapRow = 1,
                          headerStyle = NULL,
                          borders = "columns",
                          freezePane_args = list(),
                          writeData_args = list()
                          ){
  
  ls_ls_df <- rlang::list2(...)
  sheet_names <- names(ls_ls_df)
  
  fp_args <- rep_list_elements(freezePane_args, length.out = length(ls_ls_df))
  
  wb <- openxlsx::createWorkbook()
  
  # Iterate over Tabs (Worksheets)
  for (i in seq_along(ls_ls_df)) {
    
    # Take list of DF out
    ls_df <- ls_ls_df[[i]]
    # Freez pane args at each tabs
    fp_args_tab <- lapply(fp_args, `[`, i)
    
    # Header Style
    if (is.null(headerStyle)) {
        # Default HS
        hs <- createHeaderStyles(n = length(ls_df))
    }else{
        hs <- headerStyle
    }
    # Add Tabs
    openxlsx::addWorksheet(wb, sheetName = sheet_names[[i]], gridLines = TRUE)
    # Write Grid Data in each tabs
    rlang::exec(
      # FUN
      writeData_gridVertical,
      # Args
      wb,
      sheet = i, ls_df = ls_df,
      startRow = startRow, startCol = startCol,
      gapRow = gapRow,
      headerStyle = hs,
      borders = borders,
      !!!writeData_args
    )
    # Freez Panes
    rlang::exec(openxlsx::freezePane,
                wb, sheet = i,
                !!!fp_args_tab
                )
  }
  
  ## Save Files
  openxlsx::saveWorkbook(wb, file = file, overwrite = overwrite)
  
  return(invisible(wb))

}
```



Testing

```{r}
wb1 <- write.xlsx_gV(tab1 = ls_df12, 
              tab2 = list(iris, mtcars),
              file = path_excel("Excel_VertialGrid1.xlsx"), 
              headerStyle = NULL, 
              overwrite = TRUE, 
              freezePane_args = list(firstActiveRow = 2:3))

openXL(wb1)

```

```{r}

wb2 <- write.xlsx_gV(tab1 = ls_df12, file = path_excel("Excel_VertialGrid2.xlsx"), 
              headerStyle = list(hs1, hs2, hs1), overwrite = TRUE, 
              writeData_args = list(borderColour = "red"))

openXL(wb2)
```


```{r}
wb <- createWorkbook()

addWorksheet(wb, sheetName = "First", gridLines = TRUE)
addWorksheet(wb, sheetName = "Second", gridLines = TRUE)

writeData_gridVertical(wb, sheet = 1, ls_df = ls_df12, startRow = 1, gapRow = 1, 
                       headerStyle = list(hs1, hs2, hs2), borders = "columns")

writeData_gridVertical(wb, sheet = 2, ls_df = ls_df12, startRow = 1, gapRow = 1, 
                       headerStyle = hs1)

saveWorkbook(wb, path_excel("Excel_chLoc3.xlsx"), overwrite = TRUE)
openXL(wb)
```




### Write Data Grid Vertical

Vary: `x`, `startCol`, `startRow`, `headerStyle`


```{r writeData_gridVertical}
writeData_gridVertical <- function(wb, sheet, ls_df, 
                                   startCol = 1,
                                   startRow = 1, 
                                   gapRow = 1,
                                   headerStyle = NULL,
                                   borders = "columns",
                                   ...
                                   ) {
  
  # Vector of Start Rows
  start_rows <- get_startRow_stack(ls_df, startRow =  startRow, gapRow = gapRow)
  ## Rep Args
  args_ls_long <- rep_args_len(..., length.out = length(ls_df))
  borders <- rep(borders, length.out = length(ls_df))
  
  for (i in seq_along(ls_df)) {
    # Header Styles
    ## Check if it is list of Style objects
    is_ls_hs <- is.list(headerStyle) && purrr::every(headerStyle, ~ inherits(.x, "Style"))
      
    if(is.null(headerStyle)){
      hs <- createHeaderStyles(n = length(ls_df))[[i]]
    } else if (is_ls_hs) {
      # If supply list of header style; supply them correspondingly
      if (length(headerStyle) != length(ls_df)) stop("Length of list of headerStyle must equals to data frames.")
      hs <- headerStyle[[i]]
    } else if (inherits(headerStyle, "Style")) {
      # If not, style the same
      hs <- headerStyle
    } else {
      stop("`headerStyle` must be a 'Style' object.")
    }
    # Dots
    args_ls <- lapply(args_ls_long, `[`, i)
    
    # Execute Write !!
    rlang::exec(openxlsx::writeData,
                # Args
                wb,
                sheet = sheet, x = ls_df[[i]],
                startRow = start_rows[[i]], startCol = startCol, headerStyle = hs,
                borders = borders[[i]],
                !!!args_ls
    )
  }
  
}
```

Test: `writeData_gridVertical`

```{r}
wb <- createWorkbook()

addWorksheet(wb, sheetName = "First", gridLines = TRUE)

writeData_gridVertical(wb, sheet = 1, ls_df = ls_df12, startRow = 1, gapRow = 1, 
                       headerStyle = NULL, borders = c("rows","columns"), 
                       borderColour =c("red", "blue"))

saveWorkbook(wb, path_excel("Excel_chLoc2.xlsx"), overwrite = TRUE)
openXL(wb)

#debugonce("writeData_gridVertical")
```


#### Howto: write Data Grid Vertical

```{r}
wb <- createWorkbook()

addWorksheet(wb, sheetName = "First", gridLines = TRUE)

for (i in 1:3) {
  
  writeData(wb, sheet = 1, x = ls_df12[[i]], 
            startRow = get_startRow_stack(ls_df12, start = 1, gap = 1)[[i]],
            startCol = 1,
            headerStyle = NULL
            )
}


  freezePane(wb, sheet = 1, firstActiveRow = 5, firstActiveCol = 3)



saveWorkbook(wb, path_excel("Excel_chLoc2.xlsx"), overwrite = TRUE)
openXL(wb)
```


```{r}
hs1 %>% class()
hs1 %>% is.list()
hs1 %>% inherits("Style")
list(hs1, hs2) %>% every(~inherits(.x, "Style"))
```

 




In details:

```{r}
wb <- createWorkbook()

addWorksheet(wb, sheetName = "First", gridLines = TRUE)

writeData(wb, sheet = 1, 
          x = data.frame(Name = c("a","b", "c"), Val = c(1,2, 3)), 
          startCol = 1, startRow = 1, headerStyle = hs1)

writeData(wb, sheet = 1, x = iris, 
          startCol = 1, startRow = 5, headerStyle = hs2)


saveWorkbook(wb, path_excel("Excel_chLoc1.xlsx"), overwrite = TRUE)
openXL(wb)
```








### Get Rows Location 

Get start row location of each df to write

Input: `ls_df`, `gap`
Output: named numeric vector


```{r get_startRow_stack}
get_startRow_stack <- function(ls_df, startRow = 1, gapRow = 0) {
  
  df_nms <- names(ls_df)
  ## Height of each DF
  ht_df <- sapply(ls_df, nrow) + 1
  
  out <- numeric(length(ls_df))
  for (i in seq_along(ls_df)) {
    ## First DF
    if(i == 1L){
      out[[1]] <- startRow
      next
    }
    ## Start Row of DFn
    ### Start point + sum of height of last dfs + number of gaps
    out[[i]] <-  startRow + sum(ht_df[1:(i - 1)]) + (i - 1)*gapRow
  }
  names(out) <- df_nms
  out
}

get_startRow_stack(ls_df123)
get_startRow_stack(ls_df12, gapRow = 1)
get_startRow_stack(ls_df12, startRow = 2, gapRow = 1)
```

#### Test: get_startRow_stack()

```{r}
test_that("get_startRow_stack()",{
  expect_equal(get_startRow_stack(ls_df123), c(DF1 = 1,DF2 = 4,DF3 = 8))
})
```



#### Howto: Get Start Rows Location



```{r}
df1 <- data.frame(A = letters[1:2], B = c(1:2))
df2 <- data.frame(C = letters[3:5], D = LETTERS[3:5], E = 3:5)
df3 <- data.frame(G = letters[1:2], H = c(1:2))

ls_df123 <- list(DF1 = df1, DF2 = df2, DF3 = df3)
ls_df123
```




```{r}
start <- 1

gap <- 1
## Height of DF
ht_df <- sapply(ls_df12, nrow) + 1

## Start Row of DF1
start

## Start Row of DF2
start + ht_df[[1]] + 1*gap

## Start Row of DF3
start + sum(ht_df[1:2]) + 2*gap
```







### How To Capture Dot

```{r}
capture_dot <- function(...) {
  ls <- rlang::list2(...)
  ls
  
}

capture_dot(a = 1, b = 2, 3, NULL) 

capture_dot(i = iris, !!!list(a = Orange), mtcars) %>% names()
```

### Others
#### Try Write Vector

```{r}
l <- list(a = "A", b = "B")
v <- c(a = "A", b = "B")

wb <- createWorkbook()

addWorksheet(wb, sheetName = "First", gridLines = TRUE)


writeData(wb, sheet = 1, x = v, startCol = 1, startRow = 1)

saveWorkbook(wb, path_excel("Excel_vctr.xlsx"), overwrite = TRUE)
openXL(wb)
```


